Verifying Spatial Properties of Array Computations

### 2017-04-17 oopsla17/algorithms.tex
simplifying coalescing
Fix typo
simplified coalescing- back to the previous definition

\subsection{Infering specification automatically}
\label{subsec:inference}
%
We provide an inference procedure for generating specifications from
code, which are inserted automatically as comments. This helps support
maintenance of legacy code, and aids understanding of our
specification language. We illustrate inference using the
five-point stencil of \Cref{exm:checking}.

The preceding source code analysis converts the concrete syntax of
array subscripts into a set of subscript schemes in
$\mathcal{P}(\bz)^N$. These subscripts are then coalesced interatively
into sets of index schemes. We denote sets of index schemes by
$\mathcal{S}$, and initial set of subscript schemes from the analysis
as $\mathcal{S}_0$. For our example:
%
\begin{equation*}
\mathcal{S}_0 = \{\{0\} \times \{0\}, \{-1\} \times \{0\},
\{0\} \times \{-1\}, \{0\} \times \{1\}, \{1\} \times \{0\}\}
\end{equation*}

%For each array variable $a \in \mathsf{dom}(U)$, the algorithm
%constructs a set of $n$-dimensional rectangles covering all contiguous
%groups of schemes in $U(a)$.

We explain the algorithm in two parts. First, we show how \emph{adjacent} index
schemes are coalesced into smaller index schemes while remaining in union normal
form. Then, we explain how this smaller unioned indexing schemes are converted
to specifications.

\subsubsection{Covering}
A covering of (possibly overlapping) the subscript space
is calculated as a value of \region{N}. This is computed iteratively
by the following steps until a fixed-point is reached:

\begin{description}[leftmargin=0.5em]

\item[Coalesce intervals]
%
Index schemes which are \emph{contiguous} in one dimension are
coalesced.

\newcommand{\contig}[2]{\mathsf{contig}(#1, #2)}
\begin{defn}{[Contiguous]}
Two index schemes $S$ and $T$ are \emph{contiguous}
written $\contig{S}{T}$,
iff $\pi_i(S) = \pi_i(T)$
for all $1 \leq i \leq N$ apart from some dimension $k$ such that
$\pi_k(S) = [a, b]$ and $\pi_k(T) = [b+1, c]$.
\end{defn}
%
A coalesced set of index schemes is computed by:%
\begin{align*}
\mathsf{coalesce1}(S, \mathcal{S}) & =
 \{\,S \cup T \mid
T \in \mathcal{S} \, \wedge \,
\contig{S}{T} \} \\
\mathsf{coalesce}(\mathcal{S}) & =
\bigcup_{S \in \mathcal{S}} \mathsf{coalesce}(
\mathsf{coalesce1}(S, \mathcal{S})
\cup (\mathcal{S} - \{S\})) \\
\mathsf{coalesceTop}(\mathcal{S}) & =
 \mathsf{coalesce}(\mathcal{S})
\; \cup \;
\{ S \in \mathcal{S} \mid \neg \exists T \in \mathcal{S} . \; \contig{S}{T}\}
\end{align*}
\dnote{May need to prove that this is total}
%\begin{align*}
%\mathsf{coalesce}(\mathcal{S}) = \; &
%\{\,S \cup T \mid
%\forall S \in \mathcal{S}, \;
%\forall T \in \mathcal{S}. \;
%\contig{S}{T} \} \\
%\cup \; &
%\{ \,\;\;\; S \;\;\; \mid \forall S \in \mathcal{S},
%\; \not \exists T \in \mathcal{S} . \; \contig{S}{T}\}
%\end{align*}
%
That is, the new set of index schemes $\mathsf{coalesce}(\mathcal{S})$
comprises the union of the set of coalesced pairs of contiguous
schemes and the set of index schemes for which there is no other
contiguous scheme.

\item[Remove redundant schemes]
Any indexing schemes which are subsumed by another are filtered out:
%
\begin{align*}
\textsf{nonOverlap}(\mathcal{S}) =
\{S \mid \forall S \in \mathcal{S}, \; \neg \exists T \in (\mathcal{S} -
  \{S\}), \; \forall i . \; \pi_i(S) \subseteq \pi_i(T) \}
\end{align*}
%
\dnote{I'm starting to suspect that this step is actually redundant
  now- but I will need to do some math to check}
%\item \textbf{Covering}
\end{description}

The least fixed-point of $\textsf{nonOverlap} \circ \textsf{coalesce}$
is computed, providing a minimal indexing scheme covering the space.
\dnote{Proof?}

For our example:
%
\begin{align*}
\textsf{coalesce}(\mathcal{S}_0) & =
\{[-1,0] \times [0,0],\,[0,0] \times [-1, 0],\,[0,0] \times [0,1],\,[0,1]
  \times [0,0]\} \\
%\textsf{nonOverlap}(\textsf{coalesce}(\mathcal{S}_0)) & =
%                                                        \textsf{coalesce}(\mathcal{S}_0)
%  \\
\textsf{coalesce}^2(\mathcal{S}_0) & =
\{[-1, 1] \times [0, 0],\,[0, 0] \times [-1, 1]\}
\end{align*}
%
The fixed-pointed is reached within two applications of
\textsf{coalesce} for our example.
\dnote{We probably need to say something about classifying index
schemes are \region{N}?}

>>>>

\subsection{Infering specification automatically}
\label{subsec:inference}
%
\noindent
We provide an inference procedure for generating specifications from
code, which are inserted automatically as comments. This helps support
maintenance of legacy code, and aids understanding of our
specification language. We illustrate inference using the
five-point stencil of \Cref{exm:checking}.

The preceding source code analysis converts the concrete syntax of
array subscripts into a set of subscript schemes in
$\mathcal{P}(\bz)^N$. These subscripts are then coalesced iteratively
into sets of index schemes. We denote sets of index schemes by
$\mathcal{S}$, and initial set of subscript schemes from the analysis
as $\mathcal{S}_0$. For our example:
%
\begin{equation*}
\mathcal{S}_0 = \{\{0\} \times \{0\}, \{-1\} \times \{0\},
\{0\} \times \{-1\}, \{0\} \times \{1\}, \{1\} \times \{0\}\}
\end{equation*}
%
%For each array variable $a \in \mathsf{dom}(U)$, the algorithm
%constructs a set of $n$-dimensional rectangles covering all contiguous
%groups of schemes in $U(a)$.
%
We explain the algorithm in two parts. First, we show how
\emph{contiguous} index schemes are coalesced into a smaller set of
index schemes while remaining in union normal form. Then, we explain
how this smaller unioned indexing schemes are converted to
specifications.

\subsubsection{Covering}
A covering of (possibly overlapping) intervals is calculated as a
value of \region{N} by coalescing \emph{contiguous} index schemes
until a fixed-point is reached.

\newcommand{\contig}[2]{\mathsf{contig}(#1, #2)}
\begin{defn}[Contiguity]
  \label{def:contiguity}
  Two index schemes $S$ and $T$ are \emph{contiguous} written
  $\contig{S}{T}$, iff $\pi_i(S) = \pi_i(T)$ for all $1 \leq i \leq N$
  apart from some dimension $k$ such that $\pi_k(S) = [a, b]$ and
  $\pi_k(T) = [b+1, c]$.
\end{defn}
%
A coalesced set of index schemes is computed by:%
%
\begin{align*}
  \mathsf{coalesce}(\mathcal{S}) = \; & \;\;
  \{\,S \cup T \mid S \in \mathcal{S}, T \in \mathcal{S} \, \wedge \,
                                        \contig{S}{T} \}
  \;\;\; \cup \;\;\;
  \{\,S \in \mathcal{S} \mid \neg \exists T \in \mathcal{S} . \; \contig{S}{T}\}
\end{align*}
%
Thus, $\mathsf{coalesce}(\mathcal{S})$
comprises the union of the set of coalesced pairs of contiguous
schemes and the set of index schemes for which there is no other
contiguous scheme.
%
\begin{lemma}[Closure]
  For a set of index schemes $\mathcal{S}$, then
  $\mathsf{coalesce}(\mathcal{S})$ is well-defined and is a set of
  index schemes.
\end{lemma}
\begin{proof}
  By \Cref{lem:vector-union}, the union of two index schemes is an
  index scheme if all components are equal except at most one
  component. Contiguity (\Cref{def:contiguity}) matches
  this precondition, thus the first part of the union in
  \textsf{coalesce} produces a set of index schemes. The second part
  comprises a subset of the original index schemes.
\end{proof}
%
For our example, the fixed-pointed of \textsf{coalesce} is reached within two steps:
%
\begin{align*}
  \textsf{coalesce}(\mathcal{S}_0) & =
  \{[-1,0] \times [0,0],\,[0,0] \times [-1, 0],\,[0,0] \times
    [0,1],\,[0,1] \times [0,0]\} \\
  \textsf{coalesce}^2(\mathcal{S}_0) & =
  \{[-1, 1] \times [0, 0],\,[0, 0] \times [-1, 1]\} = \textsf{coalesce}^3(\mathcal{S}_0)
\end{align*}
%
\todo{Prove that this is in $Region_N$}
